---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r echo=FALSE}
# set some defaults
options(digits = 4) # limit numbers to 4 decimal places
options(scipen = 1) # don't express numbers in scientific notation
```

## Loading and preprocessing the data

```{r}
# extract zip file and load in resulting CSV
unzip("activity.zip")
activity <- read.csv("activity.csv")
# normalize the interval (pad to 4 characters)
activity$interval <- sprintf("%04d", activity$interval)
# cast $date to Date type
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?
```{r}
# subset only measured observations
activity_measured <- subset(activity, !is.na(activity$steps))
# split up the steps by date and sum each group
activity_sum <- sapply(split(activity_measured$steps, activity_measured$date), sum)
# create histogram
hist(activity_sum, breaks=20, xlab="Steps Per Day",
     main="Histogram - Steps Per Day (Oct. & Nov.)")
```

The mean of the total number of steps taken per day is `r mean(activity_sum)`  
The media of the total number of steps taken per day is `r median(activity_sum)`


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
